Jahn-Teller effect with rigid octahedral rotations in perovskites

### 2023-09-15 manuscript/paper.tex
full lattice vs AB. abstract.

\title{Cooperative Jahn-Teller effect with rigid octahedral rotations in perovskites}

@@

@@

AB ions: cubic and octahedral cages....
O ions: octahedral cage....
Cubic structure: octahedral fields of AB and O are aligned... oriented 
Tilting and rotation of the O octahedra.... change their relative orientation....
and the the combined field is no more octahedral but has a lower symmetry. 
AB lattice: V: -10 to 17 percent of V of O lattice....
So... it is a strong perturbation to the octahedral field... and couples eg and t2g manifolds... and in lifts their degeneracies....

\end{abstract}
%\maketitle

@@

From the above results, we also see that
a comparison between the rotation induced splittings
 in the two
cases (FL vs AB cages)
at a fixed value of $q_A$
would show an underestimation or an overestimation 
depending on whether $q_A$ lies to the right or the left side of the 
middle point of the two crossings.
The character of the eigenstates, however,
would still be is disagreement at $q_A$
between the two crossings.

>>>>

\title{Jahn-Teller effect with rigid octahedral rotations in perovskites} %Cooperative

@@

\com{INTRO:
The octahedral rotations and tilting in ABO$_3$ perovskites
are usually explained in terms of 
%a tolerance factor that depends on
the relative sizes 
of the three types of ions involved.
}

We consider
rigid rotation and tilting of the 
%BO$_{6/2}$ 
octahedra
 in ABO$_3$ perovskites
and study the
effect of the crystal field of the nearest A and B ion cages 
%that immediately surround the octahedra
on the single electron spectrum of the 
central 
transition metal ions.
While the crystal field of the octahedra 
that 
% quenches the orbital angular momentum of the central B ion and 
creates $e_g$ and $t_{2g}$ manifolds
do not change with such rigid rotations, 
the field of 
%A and B ionic cages %that immediately surround the octahedra
A and B cages
deforms significantly %with the octahedral rotations 
from the ideal octahedral field,
leading to sizeable splittings in the $e_g$ and $t_{2g}$ manifolds of the central B ion.
%Similar to the Jahn-Teller problem,
The lowering of the ground state energy
of a given many-electron configuration
due to these splittings
can thus be an important driving force behind the octahedral rotations 
and associated structural transformations.
We find that the size of the splitting and the local orbital ordering are
determined by a competition between the A and B cages, 
as well as a competition between the tilt and rotation angles.

\com{
SUM:
While B and tilting tend to bring $xy,x^2-y^2$ down in energy,
A and rotations do the opposite.
}

\com{

@@

AB ions: cubic and octahedral cages....
O ions: octahedral cage....
Cubic structure: octahedral fields of AB and O are aligned... oriented 
Tilting and rotation of the O octahedra.... change their relative orientation....
and the the combined field is no more octahedral but has a lower symmetry. 
AB lattice: V: -10 to 17 percent of V of O lattice....
So... it is a strong perturbation to the octahedral field... and couples eg and t2g manifolds... and in lifts their degeneracies.... 
}

\end{abstract}
\maketitle

@@

From the above results, we also see that
a comparison between the rotation induced splittings
 in the two
cases (FL vs AB cages)
at a fixed value of $q_A$
would show an underestimation or an overestimation 
depending on whether $q_A$ lies to the right or the left side of the 
middle point of the two crossings.
Besides, the characters of the eigenstates
would be in a disagreement at $q_A$
between the two crossings.
